---
title: 'Building an Automated Crypto Trader Part 2: Automating Crypto Price Collection with R and Azure'
author: ''
date: '2026-02-11'
slug: []
categories:
  - Deep Dives
tags:
  - Algo-Trading
  - R
  - Crypto
  - Azure
  - Cloud
  - Automation
  - Data Pipeline
description: "Building a production-grade automated system to fetch and store Bitcoin price data from Independent Reserve using R and Azure Container Instances"
draft: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Introduction

In my [previous post on backtesting crypto trading strategies](https://sactyr.github.io/posts/2025-12-19-building-an-automated-crypto-trader-part-1-survival-of-the-fittest-backtesting/), I developed a simple moving average crossover strategy for trading Bitcoin. However, backtesting is only half the equation — to actually trade, you need **reliable, up-to-date price data**. 

Manual data collection is error-prone and unsustainable for algorithmic trading. This post walks through how I built a fully automated system to fetch hourly Bitcoin (BTC-AUD) price data from the [Independent Reserve](https://www.independentreserve.com/au) exchange and deploy it to Azure for 24/7 operation.

**What we'll cover:**

- How the price fetching mechanism works in R
- Data validation and deduplication techniques
- Deploying R code to Azure Container Instances
- Setting up automated hourly scheduling
- Cost optimization strategies

By the end, you'll have a blueprint for building your own cloud-based crypto data pipeline.

---

## Why Automate Price Collection?

For any algorithmic trading system, you need:

1. **Consistent data collection** - No gaps in your historical record
2. **Reliability** - System runs 24/7 without manual intervention  
3. **Real-time updates** - Fresh data for trading decisions
4. **Validation** - Ensure data quality before using it

Manual collection fails on all these fronts. An automated cloud-based system solves these problems while keeping costs minimal (~$6-7/month for hourly data collection).

---

## The Architecture

Here's the high-level architecture of the system:

```
┌─────────────────┐
│  Azure Logic    │──── Triggers every hour at :05
│     Apps        │
└────────┬────────┘
         │
         ▼
┌─────────────────┐
│   Container     │──── Runs R script
│   Instance      │
└────────┬────────┘
         │
         ├──────► Independent Reserve API (fetch prices)
         │
         └──────► Azure Blob Storage (read/write data)
```

**Key components:**

- **Azure Container Instances**: Runs our R script on-demand
- **Azure Blob Storage**: Stores historical price data and logs
- **Azure Logic Apps**: Triggers the container every hour
- **Managed Identity**: Secure authentication (no hardcoded credentials)

---

## Part 1: The R Price Fetching Mechanism

### Connecting to Independent Reserve API

Independent Reserve provides a public API for historical trade data. Here's the core function:

```{r}
#' Get hourly historical price summary from Independent Reserve
get_price_history <- function(
    pri_curr_code
    ,sec_curr_code
    ,number_of_past_hours = 240
    ,ir_pub_api_url
) {
  
  url <- file.path(ir_pub_api_url, "GetTradeHistorySummary")
  
  # Build the request
  req <- request(url) %>% 
    req_url_query(
      primaryCurrencyCode = pri_curr_code
      ,secondaryCurrencyCode = sec_curr_code
      ,numberOfHoursInThePastToRetrieve = number_of_past_hours
    ) %>% 
    req_retry(max_tries = 3)
  
  # Perform the request
  resp <- req_perform(req)
  
  # Handle response 
  if (resp_status(resp) == 200) {
    data_raw <- resp_body_json(resp)
    
    # Process into tibble format
    data <- data_raw$HistorySummaryItems %>%  
      map_dfr(as_tibble) %>% 
      mutate(
        across(
          .cols = c(StartTimestampUtc, EndTimestampUtc)
          ,.fns = ~convert_utc_aest(utc_dttm = .x)
        )
        ,dttm_updated = Sys.time()
      ) %>% 
      rename(
        sttm_aest = StartTimestampUtc
        ,edtm_aest = EndTimestampUtc
      )
    
    return(data)
  } else {
    stop(paste("Failed to retrieve price history. Status:", resp_status(resp)))
  }
}
```

**Key design decisions:**

1. **API limits**: Independent Reserve allows fetching up to 240 hours (10 days) of data per request
2. **Timezone handling**: Convert UTC timestamps to AEST immediately for consistency
3. **Retry logic**: Automatically retry up to 3 times if the API is busy
4. **Error handling**: Stop execution if the request fails (logged for debugging)

### Timezone Conversion

All timestamps from Independent Reserve come in UTC. Since I'm trading in Canberra, converting to Australian Eastern time immediately prevents timezone confusion:

```{r}
convert_utc_aest <- function(utc_dttm) {
  utc_time <- ymd_hms(utc_dttm, tz = "UTC")
  with_tz(utc_time, "Australia/Sydney")
}
```

This handles both AEST and AEDT (daylight saving) automatically.

### Data Validation and Deduplication

The trickiest part of this system is **merging new data with existing historical data** while ensuring:

- No duplicate rows
- No missing hours (gaps in data)
- Data integrity

Here's how I solve this:

**1. MD5 Hashing for Deduplication**

Each row gets a unique MD5 hash based on all its values (except the update timestamp):

```{r}
data <- data %>%
  unite(
    col = "temp_concat"
    ,-dttm_updated
    ,remove = FALSE
  ) %>%  
  mutate(md5_hash = md5(temp_concat)) %>% 
  select(-temp_concat)
```

Why MD5? If Independent Reserve ever corrects historical data (e.g., fixes a trade volume error), the hash will change and we'll detect it. This prevents both duplicates and allows updates.

**2. Smart Merging**

The merge function identifies new rows by comparing MD5 hashes:

```{r}
merge_price_history <- function(
    pri_curr_code
    ,sec_curr_code
    ,ir_pub_api_url
    ,base_df
) {
  
  # Fetch last 240 hours from API
  delta_df <- get_price_history(
    pri_curr_code = pri_curr_code
    ,sec_curr_code = sec_curr_code
    ,ir_pub_api_url = ir_pub_api_url
    ,number_of_past_hours = 240
  )
  
  # Find new rows based on MD5 hashes
  diff_rows_md5 <- setdiff(delta_df$md5_hash, base_df$md5_hash)
  
  if (length(diff_rows_md5) > 0) {
    log_info("Found ", length(diff_rows_md5), " additional price history rows")
    
    # Extract and merge new rows
    diff_df <- delta_df %>% 
      filter(md5_hash %in% diff_rows_md5)
    
    merged_df <- base_df %>% 
      bind_rows(diff_df)
    
    # Validate completeness...
    return(merged_df)
  } else {
    log_info("No new rows detected")
    return(NULL)
  }
}
```

**3. Gap Detection**

After merging, we validate that there are no missing hours:

```{r}
# Generate expected hourly sequence
missing_row <- tibble(
  sttm_aest = seq(
    min(merged_df$sttm_aest)
    ,floor_date(lubridate::now() - hours(1), unit = "hours")
    ,by = "hour"
  )
) %>%
  left_join(merged_df, by = "sttm_aest") %>% 
  filter(if_any(.cols = -sttm_aest, .fns = ~(is.na(.x))))

if (nrow(missing_row) > 1) {
  log_error("Missing rows found in price history. Requires manual investigation")
  stop()
}
```

This ensures data integrity — if there's a gap (e.g., the API was down), the script stops and sends an alert rather than silently producing incomplete data.

---

## Part 2: Azure Deployment

### Why Azure?

I chose Azure over other cloud providers for several reasons:

1. **Managed Identity**: Azure's Managed Identity is simpler than GCP's service accounts
2. **Container Instances**: Pay-per-execution model (no always-on costs)
3. **Logic Apps**: Visual workflow designer for scheduling
4. **Cost**: Running hourly costs ~$0.50/month for compute

### Prerequisites

Before deploying, you'll need:

- Azure account (free tier available)
- Docker Desktop installed
- Azure CLI installed
- R 4.5+ with required packages

### Step 1: Containerize Your R Code

Create a `Dockerfile` in your project root:

```dockerfile
# Use Rocker R base image with R 4.5.2
FROM rocker/r-ver:4.5.2

# Install system dependencies
RUN apt-get update && apt-get install -y \
    libcurl4-openssl-dev \
    libssl-dev \
    libxml2-dev \
    libsodium-dev \
    && rm -rf /var/lib/apt/lists/*

# Set working directory
WORKDIR /app

# Copy R scripts
COPY R/ /app/R/

# Install R packages
RUN R -e "install.packages(c('httr2', 'jsonlite', 'dplyr', 'tidyr', \
  'lubridate', 'logger', 'openssl', 'AzureStor', 'AzureAuth', \
  'purrr', 'readr', 'stringr'), repos='https://cran.rstudio.com/')"

# Set environment variables
ENV CRYPTO_TRADING_FOLDER=/app
ENV AZURE_CONTAINER_INSTANCE=true

# Run the script
CMD ["Rscript", "/app/R/crypto_get_price_history.R"]
```

**Key points:**

- Base image: `rocker/r-ver:4.5.2` ensures consistent R version
- System dependencies: Required for packages like `httr2` and `openssl`
- Environment variables: Tell the script it's running in Azure (not locally)

### Step 2: Azure Resource Setup

Create the necessary Azure resources:

```bash
# Login to Azure
az login

# Set variables
RESOURCE_GROUP="crypto-trading-rg"
LOCATION="australiaeast"  # Sydney data center
STORAGE_ACCOUNT="cryptobtcaud2025"  # Must be globally unique
CONTAINER_NAME="crypto-data"
SUBSCRIPTION_ID=$(az account show --query id --output tsv)

# Create resource group
az group create \
  --name $RESOURCE_GROUP \
  --location $LOCATION

# Create storage account
az storage account create \
  --name $STORAGE_ACCOUNT \
  --resource-group $RESOURCE_GROUP \
  --location $LOCATION \
  --sku Standard_LRS

# Create blob container
az storage container create \
  --name $CONTAINER_NAME \
  --account-name $STORAGE_ACCOUNT \
  --auth-mode login
```

### Step 3: Build and Push Docker Image

Create a container registry and push your image:

```bash
# Create Azure Container Registry
REGISTRY_NAME="cryptotradingregistry"

az acr create \
  --name $REGISTRY_NAME \
  --resource-group $RESOURCE_GROUP \
  --location $LOCATION \
  --sku Basic

# Enable admin access
az acr update --name $REGISTRY_NAME --admin-enabled true

# Login
az acr login --name $REGISTRY_NAME

# Build image
docker build -t $REGISTRY_NAME.azurecr.io/crypto-price-fetcher:latest .

# Push to registry
docker push $REGISTRY_NAME.azurecr.io/crypto-price-fetcher:latest
```

### Step 4: Deploy Container Instance

Deploy your containerized R code:

```bash
# Get registry password
REGISTRY_PASSWORD=$(az acr credential show \
  --name $REGISTRY_NAME \
  --query "passwords[0].value" \
  --output tsv)

# Create container instance
az container create \
  --resource-group $RESOURCE_GROUP \
  --name crypto-price-fetcher \
  --image $REGISTRY_NAME.azurecr.io/crypto-price-fetcher:latest \
  --registry-login-server $REGISTRY_NAME.azurecr.io \
  --registry-username $REGISTRY_NAME \
  --registry-password $REGISTRY_PASSWORD \
  --cpu 1 \
  --memory 1.5 \
  --os-type Linux \
  --restart-policy Never \
  --environment-variables \
    AZURE_STORAGE_ACCOUNT=$STORAGE_ACCOUNT \
    AZURE_CONTAINER_NAME=$CONTAINER_NAME \
  --assign-identity [system]
```

**Important settings:**

- `--restart-policy Never`: Container runs once and stops (we'll trigger it hourly)
- `--assign-identity [system]`: Creates a Managed Identity for secure authentication
- Environment variables: Tell the R script which Azure storage to use

### Step 5: Grant Storage Permissions

The container needs permission to read/write to Blob Storage:

```bash
# Get container's Managed Identity principal ID
PRINCIPAL_ID=$(az container show \
  --resource-group $RESOURCE_GROUP \
  --name crypto-price-fetcher \
  --query identity.principalId \
  --output tsv)

# Grant Storage Blob Data Contributor role
az role assignment create \
  --assignee $PRINCIPAL_ID \
  --role "Storage Blob Data Contributor" \
  --scope /subscriptions/$SUBSCRIPTION_ID/resourceGroups/$RESOURCE_GROUP/providers/Microsoft.Storage/storageAccounts/$STORAGE_ACCOUNT
```

### Step 6: Set Up Hourly Scheduling

Use Azure Logic Apps to trigger the container every hour:

1. Go to **Azure Portal** → Create **Logic App**
2. Choose **Consumption** plan
3. In the designer, add a **Recurrence** trigger:
   - Interval: `1`
   - Frequency: `Day`
   - Time zone: `AUS Eastern Standard Time`
   - Schedule hours: `[0,1,2,...,23]`
   - Schedule minutes: `[5]`

4. Add an **HTTP** action:
   - Method: `POST`
   - URI: `https://management.azure.com/subscriptions/{subscriptionId}/resourceGroups/{resourceGroup}/providers/Microsoft.ContainerInstance/containerGroups/crypto-price-fetcher/start?api-version=2023-05-01`
   - Authentication: `Managed Identity`
   - Audience: `https://management.azure.com/`

5. Enable the Logic App's Managed Identity and grant it permissions:

```bash
# Get Logic App principal ID
LOGIC_APP_PRINCIPAL_ID=$(az resource show \
  --resource-group $RESOURCE_GROUP \
  --name crypto-price-fetcher-scheduler \
  --resource-type Microsoft.Logic/workflows \
  --query identity.principalId \
  --output tsv)

# Grant Contributor role on container
az role assignment create \
  --assignee $LOGIC_APP_PRINCIPAL_ID \
  --role "Contributor" \
  --scope /subscriptions/$SUBSCRIPTION_ID/resourceGroups/$RESOURCE_GROUP/providers/Microsoft.ContainerInstance/containerGroups/crypto-price-fetcher
```

Now the system runs automatically every hour at 5 minutes past (00:05, 01:05, 02:05, etc.).

---

## Authentication: Local vs Cloud

One elegant aspect of this setup is how authentication works differently in local vs cloud environments:

```{r}
# Detect environment
is_azure <- Sys.getenv("AZURE_CONTAINER_INSTANCE") != "" || 
            Sys.getenv("WEBSITE_INSTANCE_ID") != ""

# Set timezone for Azure (containers default to UTC)
if (is_azure) {
  Sys.setenv(TZ = "Australia/Sydney")
}

# Authenticate
get_azure_container <- function(is_azure_cloud = FALSE) {
  
  if (!is_azure_cloud) {
    # LOCAL: Use connection string from .Renviron
    conn_str <- Sys.getenv("AZURE_STORAGE_CONNECTION_STRING")
    creds <- parse_azure_connection(conn_str)
    
    blob_endpoint <- blob_endpoint(
      endpoint = creds$endpoint
      ,key = creds$account_key
    )
    
  } else {
    # AZURE CLOUD: Use Managed Identity (no credentials needed)
    token_url <- "http://169.254.169.254/metadata/identity/oauth2/token?api-version=2018-02-01&resource=https://storage.azure.com/"
    
    token_response <- httr2::request(token_url) %>%
      httr2::req_headers("Metadata" = "true") %>%
      httr2::req_perform() %>%
      httr2::resp_body_json()
    
    access_token <- token_response$access_token
    
    blob_endpoint <- blob_endpoint(
      endpoint = paste0("https://", Sys.getenv("AZURE_STORAGE_ACCOUNT"), ".blob.core.windows.net/")
      ,token = access_token
    )
  }
  
  container_name <- Sys.getenv("AZURE_CONTAINER_NAME")
  container <- storage_container(blob_endpoint, container_name)
  
  return(container)
}
```

**Local development**: Use a connection string stored in `.Renviron`

**Azure cloud**: Use the container's Managed Identity to get a token from Azure's metadata service — no secrets needed!

---

## Logging and Monitoring

### Structured Logging

The R script logs all activities:

```{r}
log_info("--- GETTING HISTORICAL PRICES START ---")
log_info("Running on ", if (is_azure) "Azure" else "Local machine")
log_success("Retrieved xbt_aud_price_history.rds [Rows: ", nrow(base_history), "]")
log_info("Found ", length(diff_rows_md5), " additional price history rows")
log_success("Price history update cycle complete.")
log_info("--- GETTING HISTORICAL PRICES END ---")
```

Logs are uploaded to Azure Blob Storage:

```
logs/
  crypto_get_price_history/
    2026-02-07/
      get_price_history_log.log
    2026-02-08/
      get_price_history_log.log
```

### Email Alerts

I've configured the Logic App to send email alerts on failures:

1. Add a **Send an email** action (Outlook/Gmail)
2. Configure it to run only when the HTTP action **fails**
3. Include error details and timestamp in AEDT

This way, I get notified immediately if something breaks.


---

## Cost Breakdown

Here's what this system actually costs per month:

| Service | Cost | Notes |
|---------|------|-------|
| Container Instances | ~$0.50 | 720 executions × ~2 seconds each |
| Blob Storage | ~$0.20 | Small data files + logs |
| Logic App | ~$0.60 | 720 workflow executions |
| Container Registry | ~$5.00 | Basic tier (stores Docker image) |
| **Total** | **~$6.30/month** | |

**Cost optimization tips:**

1. **Delete the registry between updates**: Since the container pulls the image on each start, you need to keep the registry. However, if you're not updating code frequently, you could delete it temporarily (saves $5/month), then recreate it when needed.

2. **Use Azure's free tier**: New Azure accounts get $200 credit for 30 days, which covers months of operation.

3. **Reduce execution time**: The faster your R script runs, the less you pay. Profile your code and optimize bottlenecks.

For comparison, running this on a always-on VM would cost $20-50/month. The serverless approach saves 80-90%.


---

## Lessons Learned

### 1. Timezone Headaches

Initially, I didn't set the timezone in Azure containers, so all logs showed UTC timestamps. Adding this one line fixed it:

```{r}
if (is_azure) Sys.setenv(TZ = "Australia/Sydney")
```

Now all timestamps are in AEDT/AEST, making debugging much easier.

### 2. MD5 Hashing is Essential

My first version used simple timestamp-based deduplication. This failed when Independent Reserve corrected historical data (they occasionally fix erroneous trades). MD5 hashing detects both duplicates AND changes to existing rows.

### 3. Managed Identity Propagation

Azure role assignments can take 1-2 minutes to propagate. After granting permissions, wait before testing, or you'll get mysterious 403 errors.

---

## Next Steps

This price collection system is the foundation for my trading bot. The next steps are:

1. **Build the trading execution script** - Use the same Azure architecture to execute trades based on signals
2. **Implement monitoring dashboards** - Azure Application Insights for real-time metrics
3. **Add data quality checks** - Detect anomalies (e.g., price spikes due to API errors)

The beauty of this architecture is that it's **modular**—I can add new containers for different tasks (backtesting, signal generation, trade execution) all communicating via Blob Storage.

---

## Conclusion

Building an automated crypto data pipeline taught me several valuable lessons:

- **Cloud-native R is viable**: With containers, R works great in production
- **Serverless saves money**: Pay-per-execution beats always-on VMs
- **Data validation is critical**: Never trust API data blindly
- **Azure Managed Identity is elegant**: No secrets management needed

The entire system costs less than a Netflix subscription while providing reliable, validated hourly data for algorithmic trading.

If you're building similar systems, I hope this guide saves you the trial-and-error I went through. The full code is available on [GitHub](https://github.com/sactyr/crypto_trading) (coming soon).

**Have questions or suggestions?** Leave a comment or reach out on [LinkedIn](https://linkedin.com/in/nagakaruppiah).
